Study of activity monitoring data
=================================

## Loading and preprocessing the data.

```{r,results="hide"}
setwd("D:\\Mes Documents\\Rscript")
x<-read.csv(".\\activity.csv",header=TRUE, sep=",",dec=".", na.strings = "NA")
x$date<-as.Date(x$date, "%Y-%m-%d")
```
## What is mean total number of steps taken per day?
First, we plot the total number of steps by day, for all interval
```{r}
y<-aggregate(x$steps , list(date=x$date), sum)
names(y)<- c("date","steps")
with(y,plot(date,steps,type="h", xlab="date", ylab="Total number of steps"))
```

Then, we calculate the mean and median value of steps for each date.
```{r sowtable, results='asis'}
z1<-aggregate(x$steps , list(date=x$date), mean, na.rm=TRUE)
names(z1)<- c("date","Mean steps")
z2<-aggregate(x$steps , list(date=x$date), median, na.rm=TRUE)
names(z2)<- c("date","Median steps")
z<-merge(z1,z2,by.x="date",by.y="date")
z$date<-as.character(z$date)
library(xtable)
xt<-xtable(z[,c(1:3)])
print(xt,type="html")
```
## What is the average daily activity pattern?
I make a histogram of the total number of steps taken each day :
```{r}
t<-aggregate(x$steps , list(date=x$interval), mean, na.rm=TRUE)
names(t)<- c("interval","steps")
with(t,plot(interval,steps,type="l", xlab="Interval", ylab="Average number of steps"))
```

Then I calculate the steps for which the mean number of steps is maximum.
```{r}
t[t$steps==max(t$steps),][,c(1)]
```
## Imputing missing values

The total number of NA rows :
```{r}
sum(is.na(x$steps))
```

I fill the NA with the mean value for the day.

```{r}
x_fill<-merge(x,t,by.x="interval",by.y="interval")
names(x_fill)<-c("interval","steps","date","mean_steps")
for(i in 1:nrow(x_fill)) {
  if (is.na(x_fill[i,2]))  x_fill[i,2]<- x_fill[i,4]
} 
x_fill<-x_fill[,c(1:3)]
```

Then, I calculate the mean and median value of steps for each date with ten filled data.

```{r results='asis'}
z1<-aggregate(x_fill$steps , list(date=x_fill$date), mean, na.rm=TRUE)
names(z1)<- c("date","Mean steps")
z2<-aggregate(x_fill$steps , list(date=x_fill$date), median, na.rm=TRUE)
names(z2)<- c("date","Median steps")
z<-merge(z1,z2,by.x="date",by.y="date")
z$date<-as.character(z$date)
library(xtable)
xt<-xtable(z[,c(1:3)])
print(xt,type="html")
```

## Are there differences in activity patterns between weekdays and weekends?
I calculate the type of the day in a new column.
```{r}
x_fill$typeDay <- "weekday"
for(i in 1:nrow(x_fill)) {
  if (weekdays(x_fill[i,3])=="samedi"||weekdays(x_fill[i,3])=="dimanche")  x_fill[i,4]<- "weekend"
} 
x_fill$typeDay<-as.factor(x_fill$typeDay)
```

Then I plot the average number of steps by interval and type of the day.
```{r}
library(lattice)
t<-aggregate(x_fill$steps , list(date=x_fill$interval, typeDay=x_fill$typeDay), mean, na.rm=TRUE)
names(t)<- c("interval","typeDay","meanSteps")
xyplot(meanSteps ~ interval |typeDay, type="l", data=t, ylab="Average number of steps",layout = c(1, 2))

```

